Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-08-10 prop_letter.tex
pass over paper

\begin{abstract}

We present a mathematical framework based on quantum interval-valued
probability measures to study the effect of experimental imperfections
and finite precision measurements on defining aspects of quantum
mechanics such as contextuality and the Born rule. While foundational
results such as the Kochen-Specker and Gleason theorems fail to hold
in general in a world with limited resources, when we allow infinite
precision, their validity is recovered.  However, we are able to use
the interval-valued framework to establish bounds on the validity of
those theorems in realistic experimental environments. In this way,
not only can we quantify the idea of finite-precision measurement
within our theory, but we can also suggest a possible resolution of
the Meyer-Mermin debate on the impact of finite-precision measurement
on the Kochen-Specker theorem.

\end{abstract}

@@

\begin{comment}
\keywords{Finite precision measurement, the Born rule, Gleason's
theorem, the Kochen-Specker theorem, interval-valued probability measure,
discrete quantum theory}
\end{comment}

@@

A key aspect of a quantum measurement is its statistical nature, as reflected
by application of the Born rule to determine the probability of a given measurement 
outcome\nb\citep{Born1983,peres1995quantum,544199}. The Gleason
theorem establishes that there is no alternative to the Born rule by 
demonstrating, using reasonable continuity arguments in Hilbert spaces
of dimension $d\ge3$, the existence of a unique state 
$\rho$ consistent with the statistical predictions computed from 
that rule\nb\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
It is then legitimate to question what happens when infinite precision
cannot be achieved. We will show that, if the effect of finite precision is incorporated, 
while a QIVPM  might not be consistent
with a unique state\nb$\rho$ on all projectors defined on a Hilbert space of dimension $d\ge3$, 
there is a mathematically precise sense in which one recovers the original Gleason theorem asymptotically. 
Specifically, if for bounded resources the QIVPM is induced by some mixed states\nb$\rho'$, 
those states will converge to a unique state $\rho$ when the size of the intervals characterizing 
the uncertainties tends to zero.

The paper is organized as follows.   In Sec.\ \ref{sec:fuzzy} we begin by introducing the foundations of quantum probability space and fuzzy measurement.  We then move on to introduce the concept of quantum interval valued probability in Sec.\ \ref{sec:Interval-Uncertainty}.  This QIVPM will provide the mathematical framework to quantify the impact of finite precision measurements on quantum mechanics.  For instance, in Sec.\ \ref{sec:Kochen-Specker},  we quantify the domain of validity of a  contextual measurement, thus addressing the conditions under which the Kochen-Specker theorem applies.  This provides a way of resolving the Meyer-Mermin debate\nb\footnote{There is nothing found when google ``Meyer-Mermin debate.''\\
Cabello has referred to this as the second contextuality war~\citep{Cabello2017}.\\
Jochen Szangolies called this ``the Finite-Precision Problem''~\citep{Szangolies2015}.}, which is the main result of our Theorem \ref{cor:Kochen-Specker-IVPM}.  In Sec.\ \ref{sec:Gleason}, we study Gleason's theorem in the context of imprecise measurements, concluding that QIVPM provides a framework with quantitative bounds in which Gleason's theorem, while formally invalid in a universe with bounded resources, holds asymptotically in a mathematically precise way.  Finally, we present our conclusions in Sec.~\ref{sec:Conclusion}.

% Textbook quantum mechanics asserts that the probability of a
% measurement can be computed from wave function of the system and the
% measuring observable \emph{exactly} while real measurements must be
% finite precision.  Whether finite precision measurement in reality
% affect the idealized quantum concept, e.g.,
% contextuality\nb\cite{PhysRevLett.83.3751,Mermin1999,BarrettKent2004},
% and the very foundation of quantum mechanics is still a problem. In
% this letter, we model the finite precision quantum measurement by
% quantum interval-valued probability measures (IVPMs) which provides
% not only a uniform framework to understand Gleason's
% theorem\nb\cite{gleason1957,Redhead1987-REDINA,peres1995quantum} and the
% Kochen-Specker
% theorem\nb\cite{kochenspecker1967,peres1995quantum,Redhead1987-REDINA,Griffiths2003}
% but also some examples which cannot be explained by the Born
% rule\nb\cite{Born1983,peres1995quantum,544199,Jaeger2007}.

% The tension between finite precision measurement and the foundation of
% quantum mechanics sharpened, when the original definition of
% contextuality based on infinitely precise numbers is proved not
% experimental verifiable\nb\cite{PhysRevLett.83.3751,Kent1999}.  After a
% long debate\nb\cite{Mermin1999,Peres2003,BarrettKent2004}, people are
% still trying to update the definition of contextuality when they want
% to perform further
% experiments\nb\cite{Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}.

% We found that finite precision measurement affects not only
% contextuality but also the fundamental rule computing probability from
% a state.  In the early formulation of quantum mechanics, von Neumann
% essentially formulated a quantum state as a density matrix while
% Mackey formulated a quantum state as a quantum probability measure
% which are proved by Gleason to be the essentially same as a density
% matrix\nb\cite{Varadarajan2008}.  Combining quantum probability measures
% and the classical IVPM\nb\cite{JamisonLodwick2004} which could be used
% to formulate classical finite precision measurement, we can define
% quantum IVPMs. Surprisingly, there is a quantum IVPM which cannot be
% compatible with any density matrix according to the Born rule. Similar
% to revising the definition of contextuality because of finite
% precision measurement, the very basic definition of a quantum state or
% how to compute the probability from a state might need to be refined
% when we consider the finite precision measurement. This can also
% provide a new quantum phenomena different from the classical world,
% because IVPMs become classical so that the non-Born IVPMs disappears
% when we restricting the domain of quantum IVPMs on a commuting
% sub-event space.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fuzzy Measurements}
 \label{sec:fuzzy}
A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections of
uncertain events\nb\cite{Kolmogorov1950,544199,Griffiths2003,Grabisch2016}. In the quantum case, the
events of interest are specified by \emph{projection operators}
$P$. These include the empty projection\nb$\mathbb{0}$, the identity
projection\nb$\mathbb{1}$, projections of the form $\proj{\psi}$ where
$\ket{\psi}$ is a pure quantum state (an element of a Hilbert space),
sums of \emph{orthogonal} projections $P_0$ and $P_1$ with
$P_0P_1=\mathbb{0}$, and products of \emph{commuting} projections
$P_0$ and $P_1$ with $P_0P_1=P_1P_0$. In a well-defined quantum
probability
space\nb\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010},
each event $P_{i}$ is mapped to a probability\nb$\mu(P_{i})$ using
a probability measure\nb$\mu:\events\rightarrow[0,1]$, where $\events$
is the set of all projections on a given Hilbert space, subject to
the following constraints: $\mu(\mathbb{0})=0$, $\mu(\mathbb{1})=1$,
$\mu\left(\mathbb{1}-P\right)=1-\mu\left(P\right)$, and for each pair
of \emph{orthogonal} projections $P_{0}$ and $P_{1}$:
\begin{equation}
{\mu}\left(P_{0}+P_{1}\right)={\mu}\left(P_{0}\right)+{\mu}\left(P_{1}\right)\,.\label{eq:QuantumProbability-Addition}
\end{equation}

As an example, consider a three-dimensional Hilbert space with
orthonormal basis $\{\ket{0},\ket{1},\ket{2}\}$ and an observable
$\mathbf{O}$ with spectral decomposition
$\proj{0} + 2\,\proj{1} + 3\,\proj{2}$. Two fragments of valid
probability measures $\mu_1$ and $\mu_2$ that can associated with this
space are: $\mu_1(\proj{0}) = 0.5$, $\mu_1(\proj{1}) = 0.5$,
$\mu_1(\proj{2}) = 0$, and $\mu_2(\proj{0}) = 0.5$,
$\mu_2(\proj{1}) = 0$, $\mu_2(\proj{2}) = 0.5$. By the Born rule, the
first probability measure corresponds to the quantum system being in a
state $(\ket{0}+\alpha\ket{1})/\sqrt{2}$ and the second corresponds to
the quantum system being in the state
$(\ket{0}+\alpha\ket{2})/\sqrt{2}$, where in both cases
$|\alpha|=1$. The expectation value of the observable in the first
case is 1.5 and in the second it is 2.

@@

% In conventional quantum theory (CQT), when we measure a property
% $\mathbf{O}$ of a physical system whose state is represented by a
% density matrix\nb$\rho$, the probability to obtain the value $\lambda$
% of that property is determined by the Born rule
% $\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)=\Tr\left(\rho
%   P_{i}\right)$, where $P_{i}$ denotes the projector onto the
% eigenspace corresponding to the eigenvalue $\lambda_{i}$ of the
% Hermitian operator
% $\mathbf{O}$\nb\cite{Born1983,peres1995quantum,544199,Jaeger2007}.
% Alternatively, one may say that the state $\rho$ determines the
% probability measure $\mu_{\rho}^{\mathrm{B}}$ defined as a function
% from the set $\events$ of all possible projectors $P$ onto the
% interval $[0,1]$, and satisfies the following probability axioms\nb
% \cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010}:

% \begin{itemize}
% \item $\mu_{\rho}^{\mathrm{B}}(\mathbb{0})=0$. 
% \item $\mu_{\rho}^{\mathrm{B}}(\mathbb{1})=1$. 
% \item $\mu_{\rho}^{\mathrm{B}}\left(\mathbb{1}-P\right)=
% 1-\mu_{\rho}^{\mathrm{B}}\left(P\right)$. 
% \item For a set of mutually orthogonal projectors $\left\{ P_{i}\right\} _{i=1}^{N}$,
% $P_iP_j=\delta_{ij}P_i$,
% we have 
% \begin{equation}
% \mu_{\rho}^{\mathrm{B}}\left(\sum_{i=1}^{N}P_{i}\right)=
% \sum_{i=1}^{N}\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)
% \,.\label{eq:QuantumProbability-Addition}
% \end{equation}
% \end{itemize}
% Conversely, given any quantum probability measure\nb$\mu:\events\rightarrow[0,1]$
% satisfying the above axioms, Gleason's theorem asserts that there
% exists a unique state\nb$\rho$ such that $\mu=\mu_{\rho}^{\mathrm{B}}$
% in a Hilbert space of dimension $d\geq3$\nb\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.

% Furthermore, if we express an observable\nb$\mathbf{O}$ as its spectrum
% decomposition\nb$\sum_{i=1}^{N}\lambda_{i}P_{i}$, the expectation
% value of $\mathbf{O}$ is given by\nb\cite{peres1995quantum,544199,Jaeger2007}
% \begin{equation}
% \expval{\mathbf{O}}_{\rho}=\Tr\left(\rho\mathbf{O}\right)=\sum_{i=1}^{N}\lambda_{i}\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)\ .\label{eq:quantum-expectation}
% \end{equation}

% To formalize the idea of finite precision measurement, we next
% introduce the notion of quantum interval-valued probability measures
% (\emph{IVPMs})\nb$\bar{\mu}$. The latter replaces the infinitely precise
% interval $[0,1]$ by a finite collection $\mathscr{I}$ of
% intervals. Since the intervals $\imposs=\left[0,0\right]$ and
% $\necess=\left[1,1\right]$ should be included in $\mathscr{I}$, and
% arithmetic operations on intervals
% satisfy \begin{subequations}\label{eq:interval-operations}
% \begin{eqnarray}
%  &  & [l_{1},r_{1}]+[l_{2},r_{2}]=[l_{1}+l_{2},r_{1}+r_{2}]\textrm{ and}\\
%  &  & x[l,r]=\begin{cases}
% [xl,xr] & \textrm{for }x\ge0\,;\\{}
% [xr,xl] & \textrm{for }x\le0\,.
% \end{cases}
% \end{eqnarray}
% \end{subequations} $\bar{\mu}:\events\rightarrow\mathscr{I}$ is
% defined by the axioms: 
% \begin{itemize}
% \item $\bar{\mu}(\mathbb{0})=\left[0,0\right]$. 
% \item $\bar{\mu}(\mathbb{1})=\left[1,1\right]$. 
% \item $\bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)$. 
% \item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
% $P_{i}P_{j}=\delta_{ij}P_{i}$, we have 
% \begin{equation}
% \bar{\mu}\left(\sum_{i=1}^{N}P_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(P_{i}\right)\label{eq:QuantumInterval-valuedProbability-Inclusion}
% \end{equation}
% \end{itemize}
% condition that provides only a bound and not an equivalence as in
% Eq.~\eqref{eq:QuantumProbability-Addition}. For example, if we consider
% $\left\{ P_{1},P_{2}\right\} =\left\{ P,\mathbb{1}-P\right\} $ and
% $\bar{\mu}\left(P\right)=\left[l,r\right]$, we have, when $r-l=\ell>0$,
% \[
% \bar{\mu}(\mathbb{1})=\left[1,1\right]\subsetneq\left[1-\ell,1+\ell\right]=\bar{\mu}\left(\mathbb{1}-P\right)+\bar{\mu}\left(P\right).
% \]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Intervals of Uncertainty}
  \label{sec:Interval-Uncertainty}
We will start by reviewing IVPMs in the classical setting and then
propose our quantum generalization. In the classical setting, there
are several proposals for ``imprecise
probabilities''~\citep{Dempster1967,Shafer1976,GilboaSchmeidler1994,Marinacci1999,Weichselberger2000,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
Although these proposals differ in some details, they all share the
fact that the probability $\mu(E)$ of an event~$E$ is generalized from
a single real number to an interval\nb$[\ell,r]$
where\nb$\ell$ intuitively corresponds to the strength of evidence for
the event and~$1-r$ corresponds to the strength of evidence against the
event. The collection of intervals used in the definition of a
probability measure must satisfy certain properties that we explain
next.

First, for each interval $[\ell,r]$ we have the natural constraint
$0 \leq \ell \leq r \leq 1$ that guarantees that every element of the
interval can be interpreted as a conventional probability. We also
include $\imposs=[0,0]$ and $\necess=[1,1]$ as limiting intervals to
refer to the probability of impossible events (e.g., the null event)
and certain events (e.g., the event covering the entire sample space).
For each interval $[\ell,r]$, we also need the interval
$[1-r,1-\ell]$ so that if one interval refers to the probability of an
event, the other refers to the probability of the event's complement.
For example, if we discover as a result of an experiment that
$\mu(E) = [0.2,0.3]$ for some event~$E$, we may conclude that
$\mu(\overline{E}) = [0.7,0.8]$ for the complementary event
$\overline{E}$. In addition to these simple conditions, there are some
subtle conditions on how intervals are combined, which we discuss
next.

Let $E_1$ and $E_2$ be two disjoint events with probabilities
$\mu(E_1)=[\ell_1,r_1]$ and $\mu(E_2)=[\ell_2,r_2]$. A first attempt
at calculating the probability of the combined event that \emph{either} $E_1$
or $E_2$ occurs might be $\mu(E_1\cup E_2) = [\ell_1+\ell_2,r_1+r_2]$. In
some cases, this is indeed a sensible definition. For example, if
$\mu(E_1)=[0.1,0.2]$ and $\mu(E_2)=[0.3,0.4]$ we get
$\mu(E_1\cup E_2) = [0.4,0.6]$. But consider an event $E$ such that
$\mu(E)=[0.2,0.3]$ and hence $\mu(\overline{E})=[0.7,0.8]$. The two
events $E$ and $\overline{E}$ are disjoint; the previous naïve way
of combining the intervals would give $\mu(E\cup\overline{E})=[0.9,1.1]$
which is not a valid probability interval. Moreover the event
$E\cup\overline{E}$ is the identity event covering the entire space; its
probability interval should be $[1,1]$ which is sharper than
$[0.9,1.1]$. The problem is that the two intervals are correlated:
there is more information in the combined event than in each event
separately. In our example, even though the true probability of
$\mu(E)$ can be anywhere in the range $[0.2,0.3]$ and the true
probability of $\mu(\overline{E})$ can be in anywhere in the range
$[0.7,0.8]$, the values are not independent. Any value of
$\mu(E) \leq 0.25$ will force $\mu(\overline{E})\geq 0.75$. To account
for such subtleties, the axioms of interval-valued probability do not
use a strict equality for the combination of disjoint events. The
correct constraint regarding the probability of $E_1\cup E_2$ when $E_1$
and $E_2$ are disjoint is
$\mu(E_1\cup E_2) \subseteq [\ell_1+\ell_2,r_1+r_2]$ which, in combination
with the requirement for the identity event would force
$\mu(E\cup\overline{E})$ to always be $[1,1]$ irrespective of the
particular assignments of $\mu(E)$ or $\mu(\overline{E})$.

When combining non-disjoint events, there is a further subtlety whose
resolution will give us one final general condition. For
general, not necessarily disjoint, events $E_1$ and $E_2$, we have:
\begin{equation}
\mu(E_1\cup E_2) + \mu(E_1\cap E_2) \subseteq \mu(E_1) + \mu(E_2)
\label{eq:classicalconvex}
\end{equation}
which is a generalization of the classical inclusion-exclusion
principle that uses $\subseteq$ instead of $=$ for the same reason as
before. The new condition, known as
\emph{convexity}\nb\cite{Shapley1971,GilboaSchmeidler1994,Marinacci1999,Grabisch2016},
reduces to the previously
motivated condition when the events are disjoint, i.e., when
$\mu(E_1\cap E_2) = 0$.

We now have the necessary ingredients to define QIVPMs as a
generalization of both classical IVPMs and conventional quantum
probability measures. We will show that the quantum IVPM will reduce to
the classical IVPM when the space of quantum events is restricted to
mutually commuting sub-events.

>>>>

\begin{abstract}
We present a mathematical framework based on quantum interval-valued
probability measures to study the effect of experimental imperfections
and finite precision measurements on defining aspects of quantum
mechanics such as contextuality and the Born rule. While foundational
results such as the Kochen-Specker and Gleason theorems fail to hold
in general in a world with limited resources, when we allow infinite
precision, their validity is recovered.  However, we are able to use
the interval-valued framework to establish bounds on the validity of
those theorems in realistic experimental environments. In this way,
not only can we quantify the idea of finite-precision measurement
within our theory, but we can also suggest a possible resolution of
the Meyer-Mermin debate on the impact of finite-precision measurement
on the Kochen-Specker theorem.
\end{abstract}

@@

% \keywords{Finite precision measurement, the Born rule, Gleason's
% theorem, the Kochen-Specker theorem, interval-valued probability measure,
% discrete quantum theory}

@@

A key aspect of a quantum measurement is its statistical nature, as
reflected by application of the Born rule to determine the probability
of a given measurement
outcome\nb\citep{Born1983,peres1995quantum,544199}. The Gleason
theorem establishes that there is no alternative to the Born rule by
demonstrating, using reasonable continuity arguments in Hilbert spaces
of dimension $d\ge3$, the existence of a unique state $\rho$
consistent with the statistical predictions computed from that
rule\nb\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.  It is
then legitimate to question what happens when infinite precision
cannot be achieved. We will show that, if the effect of finite
precision is incorporated, while a QIVPM might not be consistent with
a unique state\nb$\rho$ on all projectors defined on a Hilbert space
of dimension $d\ge3$, there is a mathematically precise sense in which
one recovers the original Gleason theorem asymptotically.
Specifically, if for bounded resources the QIVPM is induced by some
states\nb$\rho'$, those states will converge to a unique state $\rho$
when the size of the intervals characterizing the uncertainties tends
to zero.

The paper is organized as follows.  In Sec.\ \ref{sec:fuzzy} we begin
by introducing the foundations of quantum probability space and fuzzy
measurement.  We then move on to introduce the concept of quantum
interval valued probability in Sec.\ \ref{sec:Interval-Uncertainty}.
This QIVPM will provide the mathematical framework to quantify the
impact of finite precision measurements on quantum mechanics.  For
instance, in Sec.\ \ref{sec:Kochen-Specker}, we quantify the domain of
validity of a contextual measurement, thus addressing the conditions
under which the Kochen-Specker theorem applies.  This provides a way
of resolving the Meyer-Mermin debate, which is the main result of our
Thm~\ref{cor:Kochen-Specker-IVPM}.  In Sec.\ \ref{sec:Gleason}, we
study Gleason's theorem in the context of imprecise measurements,
concluding that QIVPM provides a framework with quantitative bounds in
which Gleason's theorem, while formally invalid in a universe with
bounded resources, holds asymptotically in a mathematically precise
way.  Finally, we present our conclusions in
Sec.~\ref{sec:Conclusion}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fuzzy Measurements}
\label{sec:fuzzy}

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections of
uncertain
events\nb\cite{Kolmogorov1950,544199,Griffiths2003,Grabisch2016}. In
the quantum case, the events of interest are specified by
\emph{projection operators} $P$. These include the empty
projection\nb$\mathbb{0}$, the identity projection\nb$\mathbb{1}$,
projections of the form $\proj{\phi}$ where $\ket{\phi}$ is a pure
quantum state (an element of a Hilbert space), sums of
\emph{orthogonal} projections $P_0$ and $P_1$ with
$P_0P_1=\mathbb{0}$, and products of \emph{commuting} projections
$P_0$ and $P_1$ with $P_0P_1=P_1P_0$. In a well-defined quantum
probability
space\nb\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010},
each event $P_{i}$ is mapped to a probability\nb$\mu(P_{i})$ using a
probability measure\nb$\mu:\events\rightarrow[0,1]$, where $\events$
is the set of all projections on a given Hilbert space, subject to the
following constraints: $\mu(\mathbb{0})=0$, $\mu(\mathbb{1})=1$,
$\mu\left(\mathbb{1}-P\right)=1-\mu\left(P\right)$, and for each pair
of \emph{orthogonal} projections $P_{0}$ and $P_{1}$:
\begin{equation}
{\mu}\left(P_{0}+P_{1}\right)={\mu}\left(P_{0}\right)+{\mu}\left(P_{1}\right)\,.\label{eq:QuantumProbability-Addition}
\end{equation}

\noindent A quantum probability measure can be easily constructed if
one knows the current state of the quantum system by using the Born
rule.  Specifically, for each pure normalized quantum state
$\ket{\phi}$, the Born rule induces a probability measure
$\mu_{\phi}^{\mathrm{B}}$ defined as
$\mu_{\phi}^{\mathrm{B}}(P)=\ip{\phi}{P\phi}$. The situation
generalizes to mixed states
$\rho = \sum_{j=1}^{N}q_{j}\proj{\phi_{j}}$, where
$\sum_{j=1}^{N}q_{j}=1$ in which case the generalized Born rule
induces a probability measure $\mu_{\rho}^{\mathrm{B}}$ defined as
$\mu_{\rho}^{\mathrm{B}}\left(P\right) = \Tr\left(\rho P\right) =
\sum_{j=1}^{N}
q_{j}\mu_{\phi_{j}}^{\mathrm{B}}\left(P\right)$~\cite{peres1995quantum,544199,Jaeger2007}.

As an example, consider a three-dimensional Hilbert space with
orthonormal basis $\{\ket{0},\ket{1},\ket{2}\}$ and an observable
$\mathbf{O}$ with spectral decomposition
$\proj{0} + 2\,\proj{1} + 3\,\proj{2}$. Two fragments of valid
probability measures $\mu_1$ and $\mu_2$ that can associated with this
space are: $\mu_1(\proj{0}) = 0.5$, $\mu_1(\proj{1}) = 0.5$,
$\mu_1(\proj{2}) = 0$, and $\mu_2(\proj{0}) = 0.5$,
$\mu_2(\proj{1}) = 0$, $\mu_2(\proj{2}) = 0.5$. By the Born rule, the
first probability measure corresponds to the quantum system being in
the pure state $(\ket{0}+\alpha\ket{1})/\sqrt{2}$ and the second
corresponds to the quantum system being in the pure state
$(\ket{0}+\alpha\ket{2})/\sqrt{2}$, where in both cases
$|\alpha|=1$. The expectation value of the observable in the first
case is 1.5 and in the second it is 2.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Intervals of Uncertainty}
\label{sec:Interval-Uncertainty}

We will start by reviewing IVPMs in the classical setting and then
propose our quantum generalization. In the classical setting, there
are several proposals for ``imprecise
probabilities''~\citep{Dempster1967,Shafer1976,GilboaSchmeidler1994,Marinacci1999,Weichselberger2000,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
Although these proposals differ in some details, they all share the
fact that the probability $\mu(E)$ of an event~$E$ is generalized from
a single \emph{real number} to an \emph{interval}\nb$[\ell,r]$
where\nb$\ell$ intuitively corresponds to the strength of evidence for
the event~$E$ and~$1-r$ corresponds to the strength of evidence
against the same event. The collection of intervals used in the
definition of a probability measure must satisfy certain properties
that we explain next.

First, for each interval $[\ell,r]$ we have the natural constraint
$0 \leq \ell \leq r \leq 1$ that guarantees that every element of the
interval can be interpreted as a conventional probability. We also
include $\imposs=[0,0]$ and $\necess=[1,1]$ as limiting intervals to
refer to the probability of impossible events (e.g., the null event)
and events that are certain (e.g., the event covering the entire
sample space).  For each interval $[\ell,r]$, we also need the
interval $[1-r,1-\ell]$ so that if one interval refers to the
probability of an event, the other refers to the probability of the
event's complement.  For example, if we discover as a result of an
experiment that $\mu(E) = [0.2,0.3]$ for some event~$E$, we may
conclude that $\mu(\overline{E}) = [0.7,0.8]$ for the complement event
$\overline{E}$. In addition to these simple conditions, there are some
subtle conditions on how intervals are combined, which we discuss
next.

Let $E_1$ and $E_2$ be two disjoint events with probabilities
$\mu(E_1)=[\ell_1,r_1]$ and $\mu(E_2)=[\ell_2,r_2]$. A first attempt
at calculating the probability of the combined event that
\emph{either} $E_1$ or $E_2$ occurs might be
$\mu(E_1\cup E_2) = [\ell_1+\ell_2,r_1+r_2]$. In some cases, this is
indeed a sensible definition. For example, if $\mu(E_1)=[0.1,0.2]$ and
$\mu(E_2)=[0.3,0.4]$ we get $\mu(E_1\cup E_2) = [0.4,0.6]$. But
consider an event $E$ such that $\mu(E)=[0.2,0.3]$ and hence
$\mu(\overline{E})=[0.7,0.8]$. The two events $E$ and $\overline{E}$
are disjoint; the previous naïve way of combining the intervals would
give $\mu(E\cup\overline{E})=[0.9,1.1]$ which is not a valid
probability interval. Moreover the event $E\cup\overline{E}$ is the
identity event covering the entire space; its probability interval
should be $[1,1]$ which is sharper than $[0.9,1.1]$. The problem is
that the two intervals are correlated: there is more information in
the combined event than in each event separately. In our example, even
though the true probability of $\mu(E)$ can be anywhere in the range
$[0.2,0.3]$ and the true probability of $\mu(\overline{E})$ can be in
anywhere in the range $[0.7,0.8]$, the values are not independent. Any
value of $\mu(E) \leq 0.25$ will force $\mu(\overline{E})\geq
0.75$. To account for such subtleties, the axioms of interval-valued
probability do not use a strict equality for the combination of
disjoint events. The correct constraint regarding the probability of
$E_1\cup E_2$ when $E_1$ and $E_2$ are disjoint is
$\mu(E_1\cup E_2) \subseteq [\ell_1+\ell_2,r_1+r_2]$ which, in
combination with the requirement for the identity event would force
$\mu(E \cup \overline{E})$ to always be $[1,1]$ irrespective of the
particular assignments of $\mu(E)$ or $\mu(\overline{E})$.

When combining non-disjoint events, there is a further subtlety whose
resolution will give us the final general condition for IVPMs. For
general, not necessarily disjoint, events $E_1$ and $E_2$, we have:
\begin{equation}
\mu(E_1\cup E_2) + \mu(E_1\cap E_2) \subseteq \mu(E_1) + \mu(E_2)
\label{eq:classicalconvex}
\end{equation}
which is a generalization of the classical inclusion-exclusion
principle that uses $\subseteq$ instead of $=$ for the same reason as
before. The new condition, known as
\emph{convexity}\nb\cite{Shapley1971,GilboaSchmeidler1994,Marinacci1999,Grabisch2016},
reduces to the previously
motivated condition when the events are disjoint, i.e., when
$\mu(E_1\cap E_2) = 0$.

We now have the necessary ingredients to define the quantum extension,
QIVPMs, as a generalization of both classical IVPMs and conventional
quantum probability measures. We will show that QIVPMs reduce to
classical IVPMs when the space of quantum events is restricted to
mutually commuting sub-events, i.e., to compatible events that can be
measured simultaneously.

\begin{definition}[QIVPM]
  Assume a collection of intervals $\mathscr{I}$ including
  $\imposs=[0,0]$ and $\necess=[1,1]$ with addition and scalar
  multiplication defined as follows:
  \begin{subequations}\label{eq:interval-operations}
  \begin{eqnarray}
   &  & [\ell_{1},r_{1}]+[\ell_{2},r_{2}]=[\ell_{1}+\ell_{2},r_{1}+r_{2}]\textrm{ and}\\
   &  & x[\ell,r]=\begin{cases}
  [x\ell,xr] & \textrm{for }x\ge0\,;\\{}
  [xr,x\ell] & \textrm{for }x\le0\,.
  \end{cases}
  \end{eqnarray}
  \end{subequations}
  Then we take a QIVPM\nb$\bar{\mu}$ to be an assignment of an interval to each
  event (projection operator\nb$P$) subject to the following constraints:
  \begin{subequations}\label{eq:QIVPM-constraints}
  \begin{eqnarray}
   &  & \bar{\mu}(\mathbb{0})=\left[0,0\right]\,,\\
   &  & \bar{\mu}(\mathbb{1})=\left[1,1\right]\,,\\
   &  & \bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)\,,
  \end{eqnarray}
  \end{subequations}
  and satisfying for each pair of \emph{commuting} projections $P_0$
  and $P_1$ with $P_0P_1=P_1P_0$,
\begin{equation}
\bar{\mu}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\,.
\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{definition}
\noindent The first three constraints, Eqs.\nb(\ref{eq:QIVPM-constraints}), are
the direct counterpart of the corresponding ones for IVPMs.  With the
understanding that the union of classical sets $E_1\cup E_2$ is
replaced by $P_0+P_1-P_0P_1$ in the case of quantum projection
operators~\cite{Griffiths2003}, the last condition,
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}), is a
direct counterpart of Eq.~(\ref{eq:classicalconvex}).

Contrary to the case with conventional quantum probability spaces,
there is generally no unique state consistent with a QIVPM. Given
observations done with finite resources, the probabilities become
intervals, and we say a state\nb$\rho$ is consistent with the
measure\nb$\bar{\mu}$ acting on a projector\nb$P$ if
$\mu_{\rho}^{\mathrm{B}}\left(P\right)\in\bar{\mu}\left(P\right)$,
where $\mu_{\rho}^{\mathrm{B}}\left(P\right)=\Tr\left(\rho P\right)$
is the probability that the event\nb$P$ happens determined by the
Born rule\nb\citep{Born1983,peres1995quantum,544199}. Given a set
of projectors\nb$S$, 
\begin{equation}
\coreBorn\left(\bar{\mu},S\right)=\set{\textrm{state }\rho}{\forall P\in S\ \mu_{\rho}^{\mathrm{B}}\left(P\right)\in\bar{\mu}\left(P\right)}
\end{equation}
is the set of all states consistent with $\bar{\mu}$ on every
projector in $S$. For example, assuming the three intervals $\imposs$,
$\necess$, and $\unknown=\left[0,1\right]$, then:
\begin{equation}
\bar{\mu}_{1}\left(P\right)=\begin{cases}
\imposs\,, & \text{if }P=\mathbb{0}\text{ ;}\\
\necess\,, & \text{if }P=\mathbb{1}\text{ ;}\\
\unknown\,, & \text{otherwise,}
\end{cases}\label{eq:no-information-IVPM}
\end{equation}
is a QIVPM. This measure represents the beliefs of an experimenter
with no prior knowledge about the particular quantum system in
question; any quantum state would be consistent with $\bar{\mu}_{1}$
on any projectors, and $\coreBorn\left(\bar{\mu}_{1},\events\right)$
is just the set of all states.

As mentioned earlier, QIVPMs become classical IVPMs when the space of
events includes only mutually commuting events. We define
$\events'\subseteq\events$ to be a sub-event space if it satisfies the
following conditions.
\begin{itemize}
\item $\left\{ \mathbb{0},\mathbb{1}\right\} \subseteq\events'$.
\item For any projection\nb$P\in\events'$, we have $\mathbb{1}-P\in\events'$. 
\item For each pair of orthogonal projections\nb$P_0 \in \events'$ and
  $P_1 \in \events'$, we have $(P_0 + P_1) \in \events'$. 
\end{itemize}
